---
title: "CIVA simulations"
author: "Jeffrey Durieux"
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(doBy))
suppressPackageStartupMessages(library(rio))
suppressPackageStartupMessages(library(plotly))
```

This report contains three simulation studies for the Clusterwise Independent Vector Analysis (C-IVA) procedure

# simulation 1

The first simulation is contains randomly generated data that is based on a CIVA model. This is a basic 'concept of proof' simulation. No special correlation structure between S across scanning sessions is imposed.

$$ X_i^{(D)} = P_{ir} S^{(D)}A_i^T$$
Note that D represents a scanning session, other terms are similar to CICA. Note that S values are generated from a multivariate Laplace distribution.

The design is as follow:

- Q components: 2, 4 or 6
- R clusters: 2 or 4
- Noise: 10%, 30% or 60% (included per D)

fixed: 

- number of voxels: 2000 
- number of scanning sessions (D): 3
- number of subjects per cluster (Nr): 20
- Time points per D: 50

Size of the data are therefore: 2000 (V) X 50 (Ti) X 3 (D) X 40 subjects (R=2) or 80 subjects (R=4)

20 repetitions were used.

Note that the Ai matrices are non-square, which is not something that you encounter in the IVA literature. This was solved by computing:

$$ A_i^{(D)} = X_i^{(D)}S^{(D)T}(S^{(D)}S^{(D)T})^{\dagger} $$

Where $\dagger$ is the Moore-Penrose pseudo inverse, this is a similar step that is done in (C-)ICA.

# results:

```{r, echo = F}
#setwd("C:/Users/jeffr/OneDrive - Erasmus University Rotterdam/Documents/CIVA/CIVA/Simulation1/")
load(file = 'allres.Rdata')

cat('Summary() of Adjusted Rand Index of clustering: \n')
summary(alldf$ari)

cat('Summary() of percentage of random starts that yielded the lowest loss function value \n')
summary(alldf$perc)

cat('Summary() of S Tucker congruence values: \n')
summary(alldf$Stuck)

cat('Summary() of A Tucker congruence values: \n')
summary(alldf$Atuck)

```

## interactions:
Congruence of S
```{r, echo=F}
summary_by(data = alldf, formula = Stuck~Err*Q*R*Nr)
```

Interactions S (means, first panel = 10% noise, middle = 30% noise, third panel 60% noise).
X axis represents components, split by clusters (colours)
```{r, echo = FALSE, warning=FALSE, message=F}
df=summary_by(data = alldf, formula = Stuck~R*Q*Nr*Err)
id1 = which(df$Err==.1)
id2 = which(df$Err==.3)
id3 = which(df$Err==.6)
p1=plot_ly(data = df[id1,], x=~jitter(Q, amount = .1), y = ~Stuck.mean,color=~as.factor(R),type = 'scatter') 
p2=plot_ly(data = df[id2,], x=~jitter(Q, amount = .1), y = ~Stuck.mean,color=~as.factor(R),type = 'scatter')
p3=plot_ly(data = df[id3,], x=~jitter(Q, amount = .1), y = ~Stuck.mean,color=~as.factor(R),type = 'scatter')
subplot(p1,p2,p3, shareY = TRUE)
```

I guess CIVA robust against noise (as was the case for CICA), difficulties with underlying model complexity (Q) but unexpected results or R effect (more clusters seems to improve results).

Congruence of A
```{r, echo=F}
summary_by(data = alldf, formula = Atuck~R*Q*Nr*Err)
```
Interactions A (means, first panel = 10% noise, middle = 30% noise, third panel 60% noise).
X axis represents components, split by clusters (colours)
```{r, echo = FALSE, warning=FALSE, message=F}
df=summary_by(data = alldf, formula = Atuck~R*Q*Nr*Err)
id1 = which(df$Err==.1)
id2 = which(df$Err==.3)
id3 = which(df$Err==.6)
p1=plot_ly(data = df[id1,], x=~jitter(Q, amount = .1), y = ~Atuck.mean,color=~as.factor(R),type = 'scatter') 
p2=plot_ly(data = df[id2,], x=~jitter(Q, amount = .1), y = ~Atuck.mean,color=~as.factor(R),type = 'scatter')
p3=plot_ly(data = df[id3,], x=~jitter(Q, amount = .1), y = ~Atuck.mean,color=~as.factor(R),type = 'scatter')
subplot(p1,p2,p3, shareY = TRUE)
```

# Simulation 2



